---
title: "R Notebook"
---


```{r}
library(openxlsx)
a <- openxlsx::read.xlsx("XS_cluster_CHARS2016 hp clustering features PCG_210915.xlsx")
b <- openxlsx::read.xlsx("XS_cluster_PCG MajorMinor for each hospital_210922.xlsx")
c <- openxlsx::read.xlsx("XS_hospital level.xlsx")

c$hospital_combine <- paste(c$hospital_level,c$hospital,sep = "_")
FT <- merge(a, b, by = "hospital", all.x = TRUE)
FT <- merge(FT, c[,c("hospital_combine","hospital_name")], by.x = "hospital", by.y = "hospital_combine", all.x = TRUE)

FT <- FT[,c(1,2,ncol(FT),3:(ncol(FT)-1))]

rm(a)
rm(b)
rm(c)

FT[is.na(FT)] <- 0

#which(colnames(FT) == "Not_grouped")
#FT <- FT[,-which(colnames(FT) == "Not_grouped")]

colnames(FT)[36]
colnames(FT)[37]
colnames(FT)[171]
colnames(FT)[172]
```


Filter out the hospitals with major procedures for trauma admissions >= 50
```{r}
library(dplyr)
FT <- FT %>% filter(num_tr_MP_sum >= 50)
```


# subgroup PCG
```{r}
library(openxlsx)
a <- openxlsx::read.xlsx("XS_cluster_CHARS2016 hp clustering features PCG_210915.xlsx")
b <- openxlsx::read.xlsx("XS_cluster_GS PCG for each hospital_210909.xlsx")
c <- openxlsx::read.xlsx("XS_hospital level.xlsx")

c$hospital_combine <- paste(c$hospital_level,c$hospital,sep = "_")
FT <- merge(a, b, by = "hospital", all.x = TRUE)
FT <- merge(FT, c[,c("hospital_combine","hospital_name")], by.x = "hospital", by.y = "hospital_combine", all.x = TRUE)
FT <- FT[,c(1,2,ncol(FT),3:(ncol(FT)-1))]

rm(a)
rm(b)
rm(c)


FT[is.na(FT)] <- 0

library(dplyr)
FT <- FT %>% filter(num_tr_MP_sum > 0)

#which(colnames(FT) == "Not_grouped")
#FT <- FT[,-which(colnames(FT) == "Not_grouped")]

colnames(FT)[3]
colnames(FT)[36]
colnames(FT)[37]

FT <- FT[,-c(3:36)]
#FT <- FT[-1,]

FT.clst <- FT[,c(3:ncol(FT))]
```



#```{r}
for (i in 35:378){
  name <- paste("p.",colnames(FT)[i],sep="")
  FT[[name]] <- FT[,i] / FT$num_tr_MP_sum
}
rm(name)

FT[is.na(FT)] <- 0
#```


Change the procedure complexity groups from raw counts to pp / MP
```{r}
FT[,c(37:171)] <- FT[,c(37:171)] / FT$num_tr_MP_sum
FT[is.na(FT)] <- 0
```


#```{r}
FT[,c(35:378)] <- FT[,c(35:378)] / max(FT[,c(35:378)])
FT[is.na(FT)] <- 0
#```

Only PCG
```{r}
FT <- FT[,c(1:3,37:171)]
```

```{r}
write.xlsx(FT, "clustering features.xlsx")
```
```{r}
FT.clst <- FT[,c(4:ncol(FT))]
```


Features for PCG in each subgroup
```{r}
library(openxlsx)
num = 34
a <- openxlsx::read.xlsx("XS_GScluster_CHARS2016 hp clustering features PCG_210804.xlsx")
b <- openxlsx::read.xlsx("XS_GScluster_PCG for each hospital_210804.xlsx")
FTsg <- merge(FT[,1:num], b, by = "hospital", all.x = TRUE)
colnames(FTsg)[(num+1):ncol(FTsg)] <- paste(colnames(FTsg)[(num+1):ncol(FTsg)],"_G",sep="")
FTsg <- merge(FTsg, a[,c("hospital","num_tr_MP_GS_sum")], by = "hospital", all.x = TRUE)
FTsg[,(num+1):(ncol(FTsg)-1)] <- FTsg[,(num+1):(ncol(FTsg)-1)] / FTsg[,ncol(FTsg)]
FTsg <- FTsg[,-ncol(FTsg)]

num <- ncol(FTsg)
print(num)
a <- openxlsx::read.xlsx("XS_Orthocluster_CHARS2016 hp clustering features PCG_210804.xlsx")
b <- openxlsx::read.xlsx("XS_Orthocluster_PCG for each hospital_210804.xlsx")
FTsg <- merge(FTsg, b, by = "hospital", all.x = TRUE)
colnames(FTsg)[(num+1):ncol(FTsg)] <- paste(colnames(FTsg)[(num+1):ncol(FTsg)],"_O",sep="")
FTsg <- merge(FTsg, a[,c("hospital","num_tr_MP_Ortho_sum")], by = "hospital", all.x = TRUE)
FTsg[,(num+1):(ncol(FTsg)-1)] <- FTsg[,(num+1):(ncol(FTsg)-1)] / FTsg[,ncol(FTsg)]
FTsg <- FTsg[,-ncol(FTsg)]

num <- ncol(FTsg)
print(num)
a <- openxlsx::read.xlsx("XS_Neucluster_CHARS2016 hp clustering features PCG_210804.xlsx")
b <- openxlsx::read.xlsx("XS_Neucluster_PCG for each hospital_210804.xlsx")
FTsg <- merge(FTsg, b, by = "hospital", all.x = TRUE)
colnames(FTsg)[(num+1):ncol(FTsg)] <- paste(colnames(FTsg)[(num+1):ncol(FTsg)],"_N",sep="")
FTsg <- merge(FTsg, a[,c("hospital","num_tr_MP_Neu_sum")], by = "hospital", all.x = TRUE)
FTsg[,(num+1):(ncol(FTsg)-1)] <- FTsg[,(num+1):(ncol(FTsg)-1)] / FTsg[,ncol(FTsg)]
FTsg <- FTsg[,-ncol(FTsg)]

num <- ncol(FTsg)
print(num)
a <- openxlsx::read.xlsx("XS_Urocluster_CHARS2016 hp clustering features PCG_210804.xlsx")
b <- openxlsx::read.xlsx("XS_Urocluster_PCG for each hospital_210804.xlsx")
FTsg <- merge(FTsg, b, by = "hospital", all.x = TRUE)
colnames(FTsg)[(num+1):ncol(FTsg)] <- paste(colnames(FTsg)[(num+1):ncol(FTsg)],"_U",sep="")
FTsg <- merge(FTsg, a[,c("hospital","num_tr_MP_Uro_sum")], by = "hospital", all.x = TRUE)
FTsg[,(num+1):(ncol(FTsg)-1)] <- FTsg[,(num+1):(ncol(FTsg)-1)] / FTsg[,ncol(FTsg)]
FTsg <- FTsg[,-ncol(FTsg)]

num <- ncol(FTsg)
print(num)
a <- openxlsx::read.xlsx("XS_Subcluster_CHARS2016 hp clustering features PCG_210804.xlsx")
b <- openxlsx::read.xlsx("XS_Subcluster_PCG for each hospital_210804.xlsx")
FTsg <- merge(FTsg, b, by = "hospital", all.x = TRUE)
colnames(FTsg)[(num+1):ncol(FTsg)] <- paste(colnames(FTsg)[(num+1):ncol(FTsg)],"_S",sep="")
FTsg <- merge(FTsg, a[,c("hospital","num_tr_MP_Sub_sum")], by = "hospital", all.x = TRUE)
FTsg[,(num+1):(ncol(FTsg)-1)] <- FTsg[,(num+1):(ncol(FTsg)-1)] / FTsg[,ncol(FTsg)]
FTsg <- FTsg[,-ncol(FTsg)]

num <- ncol(FTsg)
print(num)
a <- openxlsx::read.xlsx("XS_Othcluster_CHARS2016 hp clustering features PCG_210804.xlsx")
b <- openxlsx::read.xlsx("XS_Othcluster_PCG for each hospital_210804.xlsx")
FTsg <- merge(FTsg, b, by = "hospital", all.x = TRUE)
colnames(FTsg)[(num+1):ncol(FTsg)] <- paste(colnames(FTsg)[(num+1):ncol(FTsg)],"_Ot",sep="")
FTsg <- merge(FTsg, a[,c("hospital","num_tr_MP_Oth_sum")], by = "hospital", all.x = TRUE)
FTsg[,(num+1):(ncol(FTsg)-1)] <- FTsg[,(num+1):(ncol(FTsg)-1)] / FTsg[,ncol(FTsg)]
FTsg <- FTsg[,-ncol(FTsg)]

rm(a)
rm(b)

FTsg[is.na(FTsg)] <- 0
```

```{r}
write.xlsx(FTsg, "clustering features.xlsx")
```
```{r}
FT.clst <- FTsg[,c(4:ncol(FTsg))]
```


Feature standardization: standardize to [0, 1] range
```{r}
meanstd_stdzt <- function(x) {
  z <- (x-mean(x)) / (sd(x))
  return(z)
}
```

```{r}
minmax_stdzt <- function(x) {
  std <- (x-min(x)) / (max(x)-min(x))
  return(std)
}
```

```{r}
scalar_stdzt <- function(x) {
  z <- x / sqrt(sum(x^2))
  return(z)
  }
```

```{r}
for (i in 1:ncol(FT.clst)){
  FT.clst[,i] <- meanstd_stdzt(FT.clst[,i])
}

FT.clst[is.na(FT.clst)] <- 0
```

```{r}
for (i in 1:ncol(FT.clst)){
  FT.clst[,i] <- minmax_stdzt(FT.clst[,i])
}

FT.clst[is.na(FT.clst)] <- 0
```


```{r}
FT.clst$num_tr_adm <- minmax_stdzt(FT.clst$num_tr_adm)
FT.clst$med_tr_age <- minmax_stdzt(FT.clst$med_tr_age)
FT.clst$med_nt_age <- minmax_stdzt(FT.clst$med_nt_age)
FT.clst$med_minISS <- minmax_stdzt(FT.clst$med_minISS)
FT.clst$num_tr_MP_sum <- minmax_stdzt(FT.clst$num_tr_MP_sum)
FT.clst$num_nt_MP_sum <- minmax_stdzt(FT.clst$num_nt_MP_sum)

FT.clst[is.na(FT.clst)] <- 0
```

Standardize procedure complexity groups by min_max
```{r}
#for (i in 31:461){
#  FT.clst[,i] <- minmax_stdzt(FT.clst[,i])
#}

#FT.clst[is.na(FT.clst)] <- 0
```




PCA
```{r}
library(FactoMineR)
library(factoextra)
library(standardize)

pca <- PCA(FT.clst, scale.unit = FALSE, ncp = 40, graph = FALSE)

# Extract the score for the PCs for each individual
data <- data.frame(row.names = row.names(FT))
for (i in 1:5){
  data[,i] <- get_pca_ind(pca)$coord[,i]
}
FT.pca <- data


library(factoextra)
get_eigenvalue(pca)

fviz_eig(pca, addlabels = TRUE, ncp = 5)

# Contributions of variables to PCi
fviz_contrib(pca, choice = "var", axes = 1, top = 10)
fviz_contrib(pca, choice = "var", axes = 2, top = 10)
fviz_contrib(pca, choice = "var", axes = 3, top = 10)
fviz_contrib(pca, choice = "var", axes = 4, top = 10)
fviz_contrib(pca, choice = "var", axes = 5, top = 10)
fviz_contrib(pca, choice = "var", axes = 6, top = 10)
fviz_contrib(pca, choice = "var", axes = 7, top = 10)
fviz_contrib(pca, choice = "var", axes = 8, top = 10)
fviz_contrib(pca, choice = "var", axes = 9, top = 10)
fviz_contrib(pca, choice = "var", axes = 10, top = 10)
fviz_contrib(pca, choice = "var", axes = 11, top = 10)
fviz_contrib(pca, choice = "var", axes = 12, top = 10)
fviz_contrib(pca, choice = "var", axes = 13, top = 10)
#fviz_contrib(pca, choice = "var", axes = 14, top = 10)
#fviz_contrib(pca, choice = "var", axes = 15, top = 10)
#fviz_contrib(pca, choice = "var", axes = 16, top = 10)
#fviz_contrib(pca, choice = "var", axes = 17, top = 10)
#fviz_contrib(pca, choice = "var", axes = 18, top = 10)
#fviz_contrib(pca, choice = "var", axes = 19, top = 10)
#fviz_contrib(pca, choice = "var", axes = 20, top = 5)
#fviz_contrib(pca, choice = "var", axes = 21, top = 5)
#fviz_contrib(pca, choice = "var", axes = 22, top = 5)
#fviz_contrib(pca, choice = "var", axes = 23, top = 10)
#fviz_contrib(pca, choice = "var", axes = 24, top = 10)
#fviz_contrib(pca, choice = "var", axes = 25, top = 10)
#fviz_contrib(pca, choice = "var", axes = 26, top = 10)
#fviz_contrib(pca, choice = "var", axes = 27, top = 10)
#fviz_contrib(pca, choice = "var", axes = 28, top = 10)
#fviz_contrib(pca, choice = "var", axes = 29, top = 10)
#fviz_contrib(pca, choice = "var", axes = 30, top = 10)
#fviz_contrib(pca, choice = "var", axes = 31, top = 10)
#fviz_contrib(pca, choice = "var", axes = 32, top = 10)
#fviz_contrib(pca, choice = "var", axes = 33, top = 10)
#fviz_contrib(pca, choice = "var", axes = 34, top = 10)
#fviz_contrib(pca, choice = "var", axes = 35, top = 10)
#fviz_contrib(pca, choice = "var", axes = 36, top = 10)
#Pfviz_contrib(pca, choice = "var", axes = 37, top = 10)
#fviz_contrib(pca, choice = "var", axes = 38, top = 10)
#fviz_contrib(pca, choice = "var", axes = 39, top = 10)


# Color variables (admission type - levels) by their contribution values
fviz_pca_var(pca, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE # Avoid text overlapping
             )

# Color individuals (major ortho proc codes) by their cos2 values
# cos2 values: quality on the factor map
fviz_pca_ind(pca, col.ind = "cos2", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE # Avoid text overlapping (slow if many points)
             )

#rm(pca)
```




optimal number for clustering
```{r}
# decide the optimal k for k-means
library(factoextra)

# Elbow method
fviz_nbclust(FT.pca, kmeans, method = "wss") +
  geom_vline(xintercept = 2, linetype = 2) +
  labs(subtitle = "Elbow method")

# Silhouette method
fviz_nbclust(FT.clst, kmeans, method = "silhouette", k.max = 10)+
  labs(subtitle = "Silhouette method")
```

```{r}
unique(FT$level == FTsg$level)
```

K-means clustering
```{r}
library(xlsx)
#Cdata <- FT.pca
Cdata <- FT.clst
data <- FT.clst
level <- FT$level
kMeans <- kmeans(Cdata, 4)
#kMeans <- kmeans(Cdata, Cdata[c(1,5,2,3,7,96),])
print(kMeans$size)

# silhouette score
library(cluster)
ss <- silhouette(kMeans$cluster, dist(Cdata))
mean(ss[,3])

write.xlsx(data.frame(kMeans$cluster),"clustering result.xlsx")

# Plot the clustered data
library(Rtsne)
# Executing the algorithm on curated data
tsne <- Rtsne(data, dims = 2, perplexity = 15, verbose=TRUE, max_iter = 500, 
              check_duplicates = FALSE)
# Plotting
# Assign shapes to levels
level[which(level == "1")] <- 16
level[which(level == "2")] <- 15
level[which(level == "3")] <- 17
level[which(level == "4")] <- 8
level[which(level == "5")] <- 3
level[which(level == "non-trauma")] <- 4
level <- as.numeric(level)
# Assign colors to clustering groups
cluster <- kMeans$cluster
cluster[which(cluster == 7)] <- "lightslateblue"
cluster[which(cluster == 8)] <- "deeppink4"
cluster[which(cluster == 9)] <- "Orange"
#cluster[which(cluster == 10)] <- "Brown"
#cluster[which(cluster == 11)] <- "gold3"
#cluster[which(cluster == 12)] <- "dimgrey"

#?xyplot
library(lattice)
xyplot(tsne$Y[,2] ~ tsne$Y[,1], 
       xlab = "tSNE_x", ylab = "tSNE_y",
       main = "WA TC/non-TC clustering", 
       pch = level, col = cluster, 
       key = list(space = "right", adj=1,
                  text = list(c("level I","level II","level III",
                                "level IV","level V","non-trauma"), cex=1),
                  points = list(pch = c(16,15,17,8,3,4), col = c(1,1,1,1,1,1)),
#                  text = list(c("cluster 1","cluster 2"), cex=1),
#                  points = list(pch = 15, col = c(1,2,0,0,0,0))
                  text = list(c("cluster 1","cluster 2","cluster 3","cluster 4"), cex=1),
                  points = list(pch = 15, 
                                col = c(1,2,3,4,0,0))
#                  text = list(c("cluster 1","cluster 2","cluster 3",
#                                "cluster 4","cluster 5","cluster 6"), cex=1),
#                  points = list(pch = 15, 
#                                col = c(1,2,3,4,5,6))
                  ))

#legend("topright", legend = unique(level), pch = unique(level))
```



```{r}
library(openxlsx)
cluster1.1.2 <- openxlsx::read.xlsx("XS_result_clustering features and cluster summary_210915.xlsx", sheet = "0916_1-1-2 result")
cluster1.1.2 <- cluster1.1.2$kMeans.cluster

cluster1.2.2 <- openxlsx::read.xlsx("XS_result_clustering features and cluster summary_210915.xlsx", sheet = "0916_1-2-2 result")
cluster1.2.2 <- cluster1.2.2$kMeans.cluster

cluster1.3.2 <- openxlsx::read.xlsx("XS_result_clustering features and cluster summary_210915.xlsx", sheet = "0916_1-3-2 result")
cluster1.3.2 <- cluster1.3.2$kMeans.cluster
```

```{r}
#install.packages("fossil")
library(fossil)

rand.index(cluster1.1.2, cluster1.2.2)
rand.index(cluster1.1.2, cluster1.3.2)
rand.index(cluster1.2.2, cluster1.3.2)
#adj.rand.index(cluster1, cluster2)
```




Cluter for both continuous and categorical variables
```{r}
library(openxlsx)
FT <- openxlsx::read.xlsx("XS_result_clustering features and cluster summary_210915.xlsx", sheet = "0916_2 feature")
FT$label_GS <- as.factor(FT$label_GS)
FT$label_Ortho <- as.factor(FT$label_Ortho)
FT$label_Neu <- as.factor(FT$label_Neu)
FT$label_Uro <- as.factor(FT$label_Uro)
FT$label_Sub <- as.factor(FT$label_Sub)
FT$label_Oth <- as.factor(FT$label_Oth)

FT[is.na(FT)] <- 0

FT.clst <- FT[,c(4:ncol(FT))]
```

```{r}
library(cluster)
FT.gower <- daisy(FT.clst, metric = "gower")
```

```{r}
silhouette <- c()
silhouette = c(silhouette, NA)
for(i in 2:10){
  pam_clusters = pam(as.matrix(FT.gower),
                     diss = TRUE,
                     k = i)
  silhouette = c(silhouette, pam_clusters$silinfo$avg.width)
}

plot(1:10, silhouette, xlab = "Clusters", ylab = "Silhouette Width")
lines(1:10, silhouette)

rm(pam_clusters)
rm(silhouette)
```

```{r}
library(cluster)
pam_FT = pam(FT.gower, diss = TRUE, k = 6)
#FT[pam_german$medoids, ]

library(dplyr)
pam_summary <- FT.clst %>%
  mutate(cluster = pam_FT$clustering) %>%
  group_by(cluster) %>%
  do(cluster_summary = summary(.))
#pam_summary$cluster_summary[[1]]

library(xlsx)
write.xlsx(data.frame(pam_FT$clustering),"clustering result.xlsx")

library(plyr)
plyr::count(pam_FT$clustering)

library(Rtsne)
tsne <- Rtsne(FT.gower, dims = 2, perplexity = 15, verbose=TRUE, max_iter = 500, 
              check_duplicates = FALSE)
# Plotting
# Assign shapes to levels
level <- FT$level
level[which(level == "1")] <- 16
level[which(level == "2")] <- 15
level[which(level == "3")] <- 17
level[which(level == "4")] <- 8
level[which(level == "5")] <- 3
level[which(level == "non-trauma")] <- 4
level <- as.numeric(level)
# Assign colors to clustering groups
cluster <- pam_FT$clustering
cluster[which(cluster == 7)] <- "lightslateblue"
#cluster[which(cluster == 8)] <- "dimgrey"
#cluster[which(cluster == 9)] <- "Brown"
#cluster[which(cluster == 10)] <- "Orange"
#cluster[which(cluster == 11)] <- "gold3"
#cluster[which(cluster == 12)] <- "deeppink4"

#?xyplot
library(lattice)
xyplot(tsne$Y[,2] ~ tsne$Y[,1], 
       xlab = "tSNE_x", ylab = "tSNE_y",
       main = "WA TC/non-TC clustering", 
       pch = level, col = cluster, 
       key = list(space = "right", adj=1,
                  text = list(c("level I","level II","level III",
                                "level IV","level V","non-trauma"), cex=1),
                  points = list(pch = c(16,15,17,8,3,4), col = c(1,1,1,1,1,1)),
#                  text = list(c("cluster 1","cluster 2"), cex=1),
#                  points = list(pch = 15, col = c(1,2,0,0,0,0))
                  text = list(c("cluster 1","cluster 2","cluster 3",
                                 "cluster 4","cluster 5","cluster 6"), cex=1),
                  points = list(pch = 15, col = c(1,2,3,4,5,6))
                  ))


#library(Rtsne)
#library(ggplot2)
#tsne_object <- Rtsne(FT.gower, is_distance = TRUE)
#tsne_df <- tsne_object$Y %>%
#  data.frame() %>%
#  setNames(c("X", "Y")) %>%
#  mutate(cluster = factor(pam_FT$clustering))
#ggplot(aes(x = X, y = Y), data = tsne_df) +
#  geom_point(aes(color = cluster))
```

```{r}
a <- cbind(data.frame(pam_summary$cluster_summary[[1]]),
           data.frame(pam_summary$cluster_summary[[2]]),
           data.frame(pam_summary$cluster_summary[[3]]),
           data.frame(pam_summary$cluster_summary[[4]]),
           data.frame(pam_summary$cluster_summary[[5]]),
           data.frame(pam_summary$cluster_summary[[6]]))
```

```{r}
a <- cbind(data.frame(pam_summary$cluster_summary[[1]]),
           data.frame(pam_summary$cluster_summary[[2]]),
           data.frame(pam_summary$cluster_summary[[3]]),
           data.frame(pam_summary$cluster_summary[[4]]),
           data.frame(pam_summary$cluster_summary[[5]]),
           data.frame(pam_summary$cluster_summary[[6]]))
#write.xlsx(a,"clustering summary.xlsx")
```

```{r}
rm(pam_FT)
rm(pam_summary)
rm(tsne_df)
rm(tsne)
rm(tsne_object)
```


Test significant difference for the features across the clusters
```{r}
a <- FT$pp_nt_MP_GS
#hist(a)
#test residuals normality for ANOVA results to see if ANOVA can be used
#normality with larger p (e.g., >= 0.05)
#shapiro.test(a)
if (shapiro.test(a)$p.value >= 0.05){
  print("Y for normality")
  # Conduct ANOVA test
  #If p-value is less than the significance level 0.05: there are significant differences between the groups highlighted with "*" in the model summary.
  ANOVA <- aov(a ~ pam_FT$clustering)
  summary(ANOVA)
} else {
  print("N for normality")
  # Conduct Kruskal-Wallis test if ANOVA residuals not normal
  #If p-value is less than the significance level 0.05: there are significant differences between the treatment groups
  KW <- kruskal.test(a ~ pam_FT$clustering)
  if (KW$p.value <= 0.05){
    print("Y for sig. var. by KW")
    } else {
      print("N for sig. var. by KW")
      }
}
```



```{r}
# Test sigificant difference for categorical variables across the levels

# Conduct Chi-square Test on Homogeneity
#If p-value is less than the significance level 0.05:  there are significant differences between the treatment groups
chiS_ts <- data.frame("level" = c("1", "2", "3", "4", "5", "non-trauma"),"A" = numeric(6),"B" = numeric(6))
chiS_ts$A[[1]] <- CRTPCO_H[[1]]$num_trauma
chiS_ts$B[[1]] <- CRTPCO_H[[1]]$num_nontrauma
chiS_ts$A[[2]] <- mean(CRTPCO_H[[2]]$num_trauma)
chiS_ts$B[[2]] <- mean(CRTPCO_H[[2]]$num_nontrauma)
for (i in 3:6){
  chiS_ts$A[[i]] <- median(CRTPCO_H[[i]]$num_trauma)
  chiS_ts$B[[i]] <- median(CRTPCO_H[[i]]$num_nontrauma)
}
print(chisq.test(chiS_ts[-1]))
#print(chisq.test(chiS_ts[-1], simulate.p.value = TRUE))

```



Try tree-based model on the cluter results based on the features
```{r}
library(rpart.plot)
mod.tree <- rpart(pam_FT$clustering ~ ., FT.clst)
rpart.plot(mod.tree)
```





------------------------------------------------------------------------------------

20211005 with dummy variables and PAM

Set 1 - step 1
subgroup clustering (PAM without standardization, PCA)
```{r}
library(openxlsx)
a <- openxlsx::read.xlsx("XS_cluster_CHARS2016 hp clustering features PCG_211104.xlsx")
b <- openxlsx::read.xlsx("XS_cluster_PCG Neu for each hospital_211104.xlsx")
c <- openxlsx::read.xlsx("XS_hospital level.xlsx")

c$hospital_combine <- paste(c$hospital_level,c$hospital,sep = "_")
FT <- merge(a, b, by = "hospital", all.x = TRUE)
FT <- merge(FT, c[,c("hospital_combine","hospital_name")], by.x = "hospital", by.y = "hospital_combine", all.x = TRUE)
FT <- FT[,c(1,2,ncol(FT),3:(ncol(FT)-1))]

rm(a)
rm(b)
rm(c)

FT[is.na(FT)] <- 0

library(dplyr)
FT <- FT %>% filter(num_tr_MP_sum > 0)

colnames(FT)[3]
colnames(FT)[4]
colnames(FT)[36]
colnames(FT)[37]

FT <- FT[,-c(4:36)]

FT.clst <- FT[,c(4:ncol(FT))]
```
```{r}
file.remove("clustering features.xlsx")
write.xlsx(FT, "clustering features.xlsx")
```


Set 1 - step 2
clustering with subgroup labels and other features (PAM w standardization, PCA)
```{r}
library(openxlsx)
a <- openxlsx::read.xlsx("XS_cluster_CHARS2016 hp clustering features PCG SI_211209.xlsx")
c <- openxlsx::read.xlsx("XS_hospital level.xlsx")

c$hospital_combine <- paste(c$hospital_level,c$hospital,sep = "_")
FT <- merge(a, c[,c("hospital_combine","hospital_name")], by.x = "hospital", by.y = "hospital_combine", all.x = TRUE)
FT <- FT[,c(1,2,ncol(FT),3:(ncol(FT)-1))]

rm(a)
rm(c)

FT[is.na(FT)] <- 0

library(dplyr)
FT <- FT %>% filter(num_tr_MP_sum > 0)

file.remove("clustering features.xlsx")
write.xlsx(FT, "clustering features.xlsx")
```
(manually add labels to the feature table)
```{r}
library(openxlsx)
FT <- openxlsx::read.xlsx("XS_result_PAM clustering SI_220106.xlsx", sheet = "set1_feature")

#Feature transformation (categorical to dummy variables)
library(fastDummies)
FT <- dummy_cols(FT, select_columns = c("GS_label","Ortho_label","Neu_label",
                                        "Uro_label","Sub_label"))

FT.clst <- FT[,c(4:ncol(FT))]
FT.clst <- FT.clst[,-c(1,2,3,4,5)]
FT.clst$hp_SDI <- FT.clst$hp_SDI / 100
```
```{r}
file.remove("clustering features.xlsx")
write.xlsx(FT, "clustering features.xlsx")
```


Set 2
clustering with % PCG and other features (PAM w standardization, PCA)
```{r}
library(openxlsx)
a <- openxlsx::read.xlsx("XS_cluster_CHARS2016 hp clustering features PCG SI_211209.xlsx")
b <- openxlsx::read.xlsx("XS_cluster_PCG for each hospital_211104.xlsx")
c <- openxlsx::read.xlsx("XS_hospital level.xlsx")

c$hospital_combine <- paste(c$hospital_level,c$hospital,sep = "_")
FT <- merge(a, b, by = "hospital", all.x = TRUE)
FT <- merge(FT, c[,c("hospital_combine","hospital_name")], by.x = "hospital", by.y = "hospital_combine", all.x = TRUE)

FT <- FT[,c(1,2,ncol(FT),3:(ncol(FT)-1))]

rm(a)
rm(b)
rm(c)

FT[is.na(FT)] <- 0

colnames(FT)[39]
colnames(FT)[40]
colnames(FT)[477]
colnames(FT)[478]

library(dplyr)
FT <- FT %>% filter(num_tr_MP_sum >= 50)
FT[,c(40:477)] <- FT[,c(40:477)] / FT$num_tr_MP_sum
FT[is.na(FT)] <- 0

FT.clst <- FT[,c(4:ncol(FT))]
FT.clst$hp_SDI <- FT.clst$hp_SDI / 100
```
```{r}
file.remove("clustering features.xlsx")
write.xlsx(FT, "clustering features.xlsx")
```


Set 3
clustering with % or # Major/Minor PCG (PAM w PCA)
```{r}
library(openxlsx)
a <- openxlsx::read.xlsx("XS_cluster_CHARS2016 hp clustering features PCG_211104.xlsx")
b <- openxlsx::read.xlsx("XS_cluster_PCGMajorMinor for each hospital_211104.xlsx")
c <- openxlsx::read.xlsx("XS_hospital level.xlsx")

c$hospital_combine <- paste(c$hospital_level,c$hospital,sep = "_")
FT <- merge(a, b, by = "hospital", all.x = TRUE)
FT <- merge(FT, c[,c("hospital_combine","hospital_name")], by.x = "hospital", by.y = "hospital_combine", all.x = TRUE)

FT <- FT[,c(1,2,ncol(FT),3:(ncol(FT)-1))]

rm(a)
rm(b)
rm(c)

FT[is.na(FT)] <- 0

colnames(FT)[36]
colnames(FT)[37]
colnames(FT)[166]
colnames(FT)[167]

library(dplyr)
FT <- FT %>% filter(num_tr_MP_sum >= 50)

#FT[,c(37:166)] <- FT[,c(37:166)] / FT$num_tr_MP_sum
FT[is.na(FT)] <- 0

# Get only PCG
FT <- FT[,c(1:3,37:166)]

FT.clst <- FT[,c(4:ncol(FT))]
```

```{r}
file.remove("clustering features.xlsx")
write.xlsx(FT, "clustering features.xlsx")
```









Feature standardization
```{r}
minmax_stdzt <- function(x) {
  std <- (x-min(x)) / (max(x)-min(x))
  return(std)
}

FT.clst$num_tr_adm <- minmax_stdzt(FT.clst$num_tr_adm)
FT.clst$med_tr_age <- minmax_stdzt(FT.clst$med_tr_age)
FT.clst$med_nt_age <- minmax_stdzt(FT.clst$med_nt_age)
FT.clst$med_minISS <- minmax_stdzt(FT.clst$med_minISS)
FT.clst$num_tr_MP_sum <- minmax_stdzt(FT.clst$num_tr_MP_sum)
FT.clst$num_nt_MP_sum <- minmax_stdzt(FT.clst$num_nt_MP_sum)

FT.clst[is.na(FT.clst)] <- 0
```



PCA
```{r}
library(FactoMineR)
library(factoextra)
library(standardize)

pca <- PCA(FT.clst, scale.unit = FALSE, ncp = 40, graph = FALSE)

# Extract the score for the PCs for each individual
data <- data.frame(row.names = row.names(FT))
for (i in 1:1){
  data[,i] <- get_pca_ind(pca)$coord[,i]
}
FT.pca <- data

get_eigenvalue(pca)

fviz_eig(pca, addlabels = TRUE, ncp = 10)

#View(pca$var$contrib)

# Contributions of variables to PCi
# For a given dimension, any row/column with a contribution above the reference line could be considered as important in contributing to the dimension.
fviz_contrib(pca, choice = "var", axes = 1, top = 40)
fviz_contrib(pca, choice = "var", axes = 2, top = 30)
fviz_contrib(pca, choice = "var", axes = 3, top = 30)
fviz_contrib(pca, choice = "var", axes = 4, top = 30)
fviz_contrib(pca, choice = "var", axes = 5, top = 30)
fviz_contrib(pca, choice = "var", axes = 6, top = 35)
fviz_contrib(pca, choice = "var", axes = 7, top = 30)
fviz_contrib(pca, choice = "var", axes = 8, top = 35)
fviz_contrib(pca, choice = "var", axes = 9, top = 40)
fviz_contrib(pca, choice = "var", axes = 10, top = 20)
#fviz_contrib(pca, choice = "var", axes = 11, top = 22)
```




Find the optimal number of the clustering
```{r}
library(factoextra)
# Silhouette method
fviz_nbclust(FT.pca, cluster::pam, method = "silhouette", k.max = 10)+
  labs(subtitle = "Silhouette method")
```

```{r}
Cdata = FT.pca

library(cluster)
pam_FT = pam(Cdata, k = 10, metric = "euclidean")

library(dplyr)
pam_summary <- FT.clst %>%
  mutate(cluster = pam_FT$clustering) %>%
  group_by(cluster) %>%
  do(cluster_summary = summary(.))

# silhouette score
ss <- silhouette(pam_FT$clustering, dist(Cdata))
mean(ss[,3])

library(xlsx)
write.xlsx(data.frame(pam_FT$clustering),"clustering result.xlsx")

library(plyr)
plyr::count(pam_FT$clustering)

library(Rtsne)
tsne <- Rtsne(FT.clst, dims = 2, perplexity = 15, verbose=TRUE, max_iter = 500, 
              check_duplicates = FALSE)
# Plotting
# Assign shapes to levels
level <- FT$level
level[which(level == "1")] <- 16
level[which(level == "2")] <- 15
level[which(level == "3")] <- 17
level[which(level == "4")] <- 8
level[which(level == "5")] <- 3
level[which(level == "non-trauma")] <- 4
level <- as.numeric(level)
# Assign colors to clustering groups
cluster <- pam_FT$clustering
cluster[which(cluster == 4)] <- "blue"
cluster[which(cluster == 5)] <- "gold3"
cluster[which(cluster == 6)] <- "purple"
cluster[which(cluster == 7)] <- 5
cluster[which(cluster == 8)] <- "deeppink4"
cluster[which(cluster == 9)] <- "Orange"
cluster[which(cluster == 10)] <- "Grey"

library(lattice)
xyplot(tsne$Y[,2] ~ tsne$Y[,1], 
       xlab = "tSNE_x", ylab = "tSNE_y",
       main = "WA TC/non-TC clustering", 
       pch = level, col = cluster, 
       key = list(space = "right", adj=1,
                  text = list(c("level I","level II","level III",
                                "level IV","level V","non-trauma"), cex=1),
                  points = list(pch = c(16,15,17,8,3,4), col = c(1,1,1,1,1,1,0,0,0,0)),
                  text = list(c("cluster 1","cluster 2","cluster 3",
                                "cluster 4","cluster 5","cluster 6",
                                "cluster 7","cluster 8","cluster 9","cluster 10"), cex=1),
                  points = list(pch = 15, col = c(1,2,3,"blue","gold3","purple",5,"deeppink4","Orange","Grey"))
                  ))
```


Summary of the contributing features
```{r}
library(dplyr)
#set 1
FT.summary <- FT[,9:ncol(FT)] %>%
  mutate(cluster = pam_FT$clustering)

FT.summary <- 
  aggregate(FT.summary[,c(3,4,5,8,11,16:27,31,32,33,36:53)],
            list(FT.summary$cluster), mean)

write.xlsx(FT.summary, "clustering summary.xlsx")
```
```{r}
#set 2
FT.summary <- FT[,4:ncol(FT)] %>%
  mutate(cluster = pam_FT$clustering)

#FT.summary <-
#  aggregate(FT.summary[,c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,16,17,18,19,20,21,22,23,24,26,28,29,30,31,32,33,144,145,150,155,171,172,178,227,354,454)], list(FT.summary$cluster), mean)

FT.summary <-
  aggregate(FT.summary[,c(1:9,11:27,29,31:36,147,148,153,158,174,175,181,230,357,457)],list(FT.summary$cluster), mean)

file.remove("clustering summary.xlsx")
write.xlsx(FT.summary, "clustering summary.xlsx")
```
```{r}
#set 3-1
FT.summary <- FT[,4:ncol(FT)] %>%
  mutate(cluster = pam_FT$clustering)

FT.summary <-
  aggregate(FT.summary[,c(1,4,5,6,16,17,19,22,32,33,34,35,38,39,40,41,44,48,51,52,63,64,68,69,71,74,94,95,97,115,117,125,127,129,130)],list(FT.summary$cluster), mean)

write.xlsx(FT.summary, "clustering summary.xlsx")
```
```{r}
#set 3-2
FT.summary <- FT[,4:ncol(FT)] %>%
  mutate(cluster = pam_FT$clustering)

FT.summary <-
  aggregate(FT.summary[,c(3,16,19,25,33,38,39,40,41,44,51,69,71,95,99,105,115,117,125,129)],
            list(FT.summary$cluster), mean)

write.xlsx(FT.summary, "clustering summary.xlsx")
```



```{r}
# a: set 1 N5 without social index
a <- c(1,1,2,3,1,2,4,1,5,3,4,4,4,2,5,4,4,5,4,2,4,5,3,5,4,5,5,2,5,5,4,4,4,4,4,4,4,4,4,4,5,4,5,5,4,4,4,4,4,4,4,4,4,4,4,4,4,5,4,4,2,2,2,4,4,2,4,5,4)
b <- pam_FT$clustering
library(fossil)
rand.index(a, b)
```

```{r}
# a: set 2 N7 without social index
a <- c(1,2,2,2,2,2,3,2,4,2,4,5,4,5,5,4,4,2,5,5,4,4,5,4,2,2,5,2,6,4,4,6,6,4,6,4,4,6,4,4,4,5,4,5,4,6,2,7,2,3,2,5,5)
b <- pam_FT$clustering
library(fossil)
rand.index(a, b)
```

